\LARGE \bf Expectation-Maximization based defense mechanism for distributed Model Predictive Control

### 2022-02-11 article/article.tex
version with 6 pages, Abstr/Intro/Concl TODO

\begin{abstract}% Abstract of not more than 250 words.
  \todo[write abstract]{
  In distributed predictive control structures, communication among agents is required to achieve a consensus and approach an optimal global behavior. Such negotiation mechanisms are sensitive to attacks on these exchanges.
  This paper proposes a monitoring scheme that detects and mitigates these attacks' effects in a resource allocation framework.
  The performance of the proposed method is illustrated through simulations of the temperature control of multiple rooms under input power constraints.
}
\end{abstract}

@@

This operation is done cooperatively, where the controllers exchange information with each other.
This approach can be used, for example, for the voltage regulation of electrical networks~\citep{HansEtAl2019}, and the production-consumption balance in these same networks~\citep{HerreraEtAl2015}.
Multi-agent systems~\citep{KantamneniEtAl2015} or \dmpc{}~\citep{MaestreEtAl2014} are standard tools to design these controllers.

When one of these controllers is attacked, it can result in the violent destruction of the corrupted element (perhaps the whole system), or more subtly, it can lead to deviant behavior that is more difficult to detect.
Usually the attacks aim to de-optimize the global operation by benefiting/prejudicing some specific group of controllers, or to \todo{slowly} lead the system towards a state of malfunction/instability~\todo[use other]{\citep{NogueiraEtAl2021}}.

With this in mind, a few articles have focused on safety in \dmpc{} frameworks, presenting vulnerabilities of different frameworks, such as lagrange-based decomposition~\citep{VelardeEtAl2017}, and Jacobi-Gauss decomposition~\citep{ChanfreutEtAl2018}.

The authors in~\cite{VelardeEtAl2017} highlight the complexity of the problems. A modification of a local goal, the hacking of communication,
or the failure of one of the agents are events that disrupt global behavior.

Some strategies are presented to mitigate the effects of such event,
for instance dismissing extreme values as in~\cite{VelardeEtAl2017}, or
using secure scenarios based on reliable historical data~\cite{VelardeEtAl2017a}.
The most recent~\cite{MaestreEtAl2021} presents ways to generate these scenarios.

Our previous work~\citep{NogueiraEtAl2021} proposed a framework dedicated to
sharing resource algorithms (primal decomposition), and shown its vulnerabilities.
More precisely, \fdi{} in the communication
between a coordinator and local constraint-free agents bound by global equality constraints. The paper also proposed a safe algorithm based on data reconstruction to mitigate the effects.

@@

\todo{The paper is constructed as follows}

\emph{Notation:}
In this paper, $\norm{\cdot}$, $\norm{\cdot}_{F}$, and $\norm{\vec{v}}_{Y}$ represent the $\ell_{2}$ norm, the Frobenius norm and the weighted norm $\norm{Y^{\frac{1}{2}}\vec{v}}$.
$\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto set $\set{T}$.
% $\card{x}$ is the number of elements in $x$.
$\kron{}{}$, and $\hadamard{}{}$ are the Kronecker and Hadamard products.
$\vectorize{A}$ vectorizes matrix $A$.
${n\mathbin{:}i\mathbin{:}j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\mathbin{:}j}$ is equal to ${n\mathbin{:}1\mathbin{:}j}$.
$[A,B]$ and $[A;B]$ are horizontal and vertical concatenation of matrices.
${\0_{m}=0_{m\times 1}}$ (${\1_{m}=1_{m\times 1}}$), where
$0_{m\times n}$ ($1_{m\times n}$) is a $0$ ($1$) filled \mbox{$m$-by-$n$} matrix.
$\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
$\preceq$ is the component-wise less or equal for vectors and matrix inequality for matrices in ${\symmetric^{n}=\setbuild{X:\R^{n\times n}}{X'=X}}$.
${\defpos^{n}=\setbuild{X:\symmetric^{n}}{X\succ 0}}$, ${\semidefpos^{n}=\setbuild{X:\symmetric^{n}}{X\succeq 0}}$.
$I_{c}$ is the identity matrix of size $c$.
$A^{\dagger}$ is the generalized inverse ${{(A^{T}A)}^{-1}A^{T}}$.
$\determinant{A}$ is the determinant of $A$.
A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
$\random{a},\random{A},\randomvec{a}$ are random variables, scalar, matrix and vector respectively.
$\expectation{\random{x}}$ is the expected value of $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
% $\pdf{\cdot}$ is a probability density function.
${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.

@@

\subsection{Model Predictive Control}\label{ssec:MPC}
Let us consider a system composed of $M$ discrete-time linear time-invariant agents, whose dynamics are ruled by:
\begin{equation}
\begin{matrix}
  \label{eq:systems}
\vec{x}_{i}[k+1]&=&A_{i}\xik &+& B_{i}\uik\\
\vec{y}_{i}[k+1]&=&C_{i}\xik &&
\end{matrix},
\end{equation}
with $\vec{x}_{i}[k]:\R^{n_{x}}$, $\vec{u}_{i}[k]:\R^{n_{u}}$ and $\vec{y}_{i}[k]:\R^{n_{y}}$.

Each agent's input $\uik$ is constrained by
\begin{equation}
  \label{eq:u_local_constraints}
  \0_{n_{u}}\preceq\uik\preceq\vec{u}_{\max},
\end{equation}
with ${\umax:\R^{n_{u}}}$.
The $M$ agents are coupled by global input constraints with weighting matrices \todo{${\Gamma_{i}:\semidefpos^{n_{u}}}$}:
\begin{equation}
  \label{eq:global_constraint}
  \sum^{\nsubsystems}_{i=1}\Gamma_{i}\uik\preceq\umax.
\end{equation}

The overall system is controlled by a \mpc{}, which predicts the dynamics for a finite prediction horizon $\predhorz$ and computes the optimal input by solving the following problem:
\begin{equation}\label{eq:GOP}
\resizebox{.88\columnwidth}{!}{$
\begin{matrix}
\underset{\useq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits^{M}_{i=1} \overbrace{\sum_{j=1}^{\predhorz} \norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
\mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:u_local_constraints}\ \mathrm{and}\ \eqref{eq:global_constraint}
\left\}\small
\begin{aligned}
  &\forall i\in \{1:\nsubsystems\}\\
  &\forall j\in \{1:\predhorz\}
\end{aligned}\right.,

@@

Due to the monolithic aspect of~\eqref{eq:GOP}, solving it at each time $k$ can be quite challenging, since the prediction horizon and the number of agents can be large making the process computationally intensive. Another issue that can arise is the fact that the complete information (${\set{I}_{i}=\{A_{i},B_{i},C_{i},\vec{v}_{i}[k],Q_{i},R_{i}\}}$) from all the subsystems is needed to solve the problem,
which can be viewed negatively in a confidentiality point of view.

One can see that if it were not for the global constraints~\eqref{eq:global_constraint}, the problem would be easily decomposed in $M$ parts and solvable in parallel.
So, we use the \emph{primal decomposition} method, sometimes called \emph{quantity decomposition} or \emph{resource allocation}, with wich we can profit from the parallelism, while avoiding the use of the complete information $\set{I}_{i}$.

@@

The technique consists of decomposing the monolithic \mpc{} in~\eqref{eq:GOP} into $M$ modified \mpc{} problems~\eqref{eq:DOP_local}, called \emph{local problems}, which can be solved in parallel by each subsystem, and a \emph{master problem}~\eqref{eq:DOP_master}, which is equivalent to the global problem and is solved by a coordinator.
\begin{subequations}
  \begin{equation}
    \left.
      \small
      \begin{aligned}
        J_{i}^{\star}(\thetaik)&=\underset{\uiseq}{\mathrm{minimize}} \obji\\
        \mathrm{s.t.} &\quad\eqref{eq:systems}\ \&\ \eqref{eq:u_local_constraints}\\
        &\quad\Gamma_{i}\uik\preceq\qik:\dik\\
      \end{aligned}
    \right\}
    \small
    \begin{aligned}
      &\forall i\in \{1\mathbin{:}\nsubsystems\}\\
      &\forall j\in \{1\mathbin{:}\predhorz\}
    \end{aligned}
    \label{eq:DOP_local}
  \end{equation}
  \begin{equation}
    \small
    \begin{aligned}
      J^{\star}&=\underset{\qiseq}{\mathrm{minimize}}\sum^{\nsubsystems}_{i=1} J^{\star}_i(\qik)\\
      \mathrm{s.t.} &\quad \sum_{i=1}^{\nsubsystems}\qik\preceq\umax,\ \forall j\in\{1\mathbin{:}\predhorz\}\\
      &\quad \qik[k]\succeq\0_{n_{u}}\text{, } \forall i\in\{1\mathbin{:}M\},\forall j\in\{1\mathbin{:}\predhorz\}
    \end{aligned}
    \label{eq:DOP_master}
  \end{equation}
\end{subequations}
The modified \mpc{} problems in~\eqref{eq:DOP_local} are created by exchanging the $\vec{u}_{\max}$ in~\eqref{eq:global_constraint} by vectors $\qik$, where which one corresponds to the quantity of the total resource $\vec{u}_{\max}$ that is allocated to agent $i$ in time $[k]$ for each prediction $j$, thus the names \emph{resource allocation} and \emph{quatity decomposition}. This new set of constraints have associated dual variables ${\dik\succeq\0_{n_{u}}}$. The sequence of this variables variables ${\qiseq}$ and ${\diseq}$ can be aggregated in vectors ${\thetaik=\kron{\1_{\predhorz}}{\qik}}$, and ${\lambdaik=\kron{\1_{\predhorz}}{\dik}}$.

The \emph{master problem} finds the optimal allocation sequences $\optthetai$ by using a projected sub-gradient method:
\begin{equation}
  \label{eq:projectedSubgradient}
\vec{\theta}[k]\pplusone=\Proj^{\set{H}}(\vec{\theta}[k]\p-\rho\p\vec{g}[k]\p),
\end{equation}
where ${\set{H} = \setbuild{\vec{\theta}}{I_{c}^{M}\vec{\theta}\preceq \vec{U}_{\max}\ \&\&\ \vec{\theta}\succeq\0 }}$, being ${c=\predhorz n_{u}}$, ${I_{c}^{M}=\kron{\1_{M}}{I_{c}}}$, ${\vec{U}_{\max}=\kron{\1_{\predhorz}}{\vec{u}_{\max}}}$, $(p)$ is a given step in the iterative process and $\vec{g}\p[k]$ is a sub-gradient of the objective of the function problem in~\eqref{eq:DOP_master} in step $(p)$.

From~\cite{BoydEtAl2015}, it is known that $-\lambdaik$ are sub-gradients of objective $J[k]$.
Plugging it in $\vec{g}[k]\p$ we have
\begin{equation}
  \label{eq:thetaNegot}
\vec{\theta}[k]\pplusone=\Proj^{\set{H}}(\vec{\theta}[k]\p+\rho\p\vec{\lambda}[k]\p),
\end{equation}
\newcommand{\negotiation}{\emph{negotiation}}
which henceforth is referred as the \negotiation.

@@

\begin{algorithm2e}[h]
  \DontPrintSemicolon%
  $p:=0$\;
  Coordinator initializes $\vec{\theta}\p$ \;
  \Repeat{$\|\vec{\theta}^{(p)} -\vec{\theta}^{(p-1)}\|\leq\epsilon$}{
  Subsystems solve~\eqref{eq:DOP_local}, and send $\optlambdai (\thetai\p)$\;
  Coordinator updates allocations~\eqref{eq:thetaNegot}\;
  $p:=p+1$
}
 \caption{Quantity decomposition based \acrlong{dmpc}.}\label{alg:quantityAlg}
\end{algorithm2e}

One can observe that instead of using $\mathcal{I}_{i}$, each agent only sends $\lambdaik$.
An interpretation of $\lambdaik$ is the dissatisfaction of an agent with the resource $\thetaik$ allocated for it, where ${\lambdaik=\0_{c}}$ means total satisfaction.
The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations, but if false data is injected, the \negotiation{} can be driven by a malicious agent that can take advantage of this trust to favor itself at the expense of others or even destabilize the \negotiation{}, as shown in~\cite{NogueiraEtAl2021}.

Our main objective is to mitigate the effects of a given configuration where agents send untrustworthy of the form
\begin{equation*}
\lambdaicheat=\gamma_{i}(\lambdai).
\end{equation*}

@@

The \emph{local problems} can be rewritten in a equivalent (same solution) constrained \qp{} form:
\begin{equation}\label{eq:quadratic_case}
\begin{matrix}
\underset{\Uik}{\mathrm{minimize}}& \overbrace{\frac{1}{2} \Uik^T H_i\Uik+{\fik}^T\Uik}^{\textstyle J_{i}(\vec{\theta}_{i})}\\
\mathrm{s.t.}&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\\
&\Uik\succeq\0_{c},
\end{matrix}
\end{equation}
where ${\Uik}$ stacks the input prediction sequences, and
${\bar{\Gamma}_{i}=\kron{I_{\predhorz}}{\Gamma_{i}}}$.
The exact values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$. The approach presented in this paper does not depend on the choice of the control objective.
The approach depends only on the fact that $H_{i}:\semidefpos$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.

\no{
  We choose reference tracking just to illustrate the approach. With this objective, we have
${\vik=\wik-C_{i}\xik}$, where $\wik$ is the output reference in time $k$, which yields
\begin{equation}
\small\begin{matrix*}[l]
 H_i&=&\mathcal{D}_i^T\bar{Q}_i\mathcal{D}_i+\bar{R}_i\\
\fik&=&\mathcal{D}_i^T\bar{Q}_i(\mathcal{M}_i\xik-\Wik),
\end{matrix*}
\label{eq:matrices}
\end{equation}
where $\Wik$ stacks the setpoint prediction sequence, which is supposed to be constant during the predicted horizon ${\Wik=\kron{\1_{\predhorz}}{\vec{w}_{i}[k]}}$ (no prior information).
${\mathcal{M}_{i}}$  and ${\mathcal{D}_{i}}$ are the prediction matrices of the \mpc{}.
${\bar{Q}_{i}=\kron{I_{\predhorz}}{Q_{i}}}$, and ${\bar{R}_{i}=\kron{I_{\predhorz}}{R_{i}}}$.
}

\begin{figure}[t]
% TODO(accacio): correct image
  \centering
  \begin{tikzpicture}[node distance=1.7cm]
    \tikzset{
      agent/.style={draw,circle,minimum width=1cm,},
      coord/.style={draw,circle,minimum width=1.7cm,very thick}
    }
    % \draw[thick,blue,rounded corners=10pt] (-1.7,1.2) rectangle (1.7,-1.7);
    % \node at (1.2,-1.3) {O};
    % \node (pi) at (-2.5,.5) {$\pi$};
    % \node (phi) at (2.5,-1.) {$\phi_{c}$};

    \node[agent] (a1) {A$_{1}$};
    \node (a2) [right of=a1]  {$\dots$};
    \node[agent] (am) [right of=a2] {A$_{M}$};
    \node[coord] (c) [above of=a2] {Coord};

    \draw[-latex] (a1) to [out=60,in=210] node [sloped,above] {$\vec{\lambda_{1}}$} (c) ;
    \draw[-latex] (c) to [out=-120,in=30] node [sloped,below] {$\vec{\theta}_{1}$} (a1);

    \draw[-latex] (am) to [out=150,in=-60]node [sloped,below] {$\vec{\lambda}_{M}$} (c);
    \draw[-latex] (c) to [out=-30,in=120] node [sloped,above] {$\vec{\theta}_{M}$} (am);

  \end{tikzpicture}
  \caption{Graph representation of Coordinator exchanging with $M$ agents in a \dmpc{} framework.}\label{fig:dmpc_graph}
\end{figure}

>>>>

\begin{abstract}% Abstract of not more than 250 words.
  \todo[write abstract]{
  In distributed predictive control structures, communication among agents is required to achieve a consensus and approach an optimal global behavior. Such negotiation mechanisms are sensitive to attacks on these exchanges.
  This paper proposes a monitoring scheme based on Expectation Maximization to detect and mitigate the effects of attacks in a resource allocation framework.
  The performance is assessed through academical example of the temperature control of multiple rooms under input power constraints.
}
\end{abstract}

@@

This operation is done cooperatively, where the controllers exchange information with each other.
This approach can be used, for example, for the voltage regulation of electrical networks, and the production-consumption balance in these same networks.
Multi-agent systems~\citep{KantamneniEtAl2015} or \dmpc{}~\citep{MaestreEtAl2014} are standard tools to design these controllers.

A few articles have focused on safety in \dmpc{} frameworks, presenting vulnerabilities of different frameworks, such as Lagrange-based decomposition~\citep{VelardeEtAl2017}, Jacobi-Gauss decomposition~\citep{ChanfreutEtAl2018} and primal decomposition~\citep{NogueiraEtAl2021}.

The authors in these works highlight the complexity of the problems. A modification of a local goal, the hacking of communication, or the failure of one of the agents are events that disrupt global behavior.
If one of these agents is attacked, it can result in the violent destruction of the corrupted element (perhaps the whole system), or more subtly, it can lead to deviant behavior that is more difficult to detect.
% Usually the attacks aim to de-optimize the global operation by benefiting/prejudicing some specific group of controllers, or to lead the system towards a state of malfunction/instability~\todo[use other]{}.

Some strategies are presented to mitigate the effects of such event.
For instance dismissing extreme values as in~\cite{VelardeEtAl2017}, or
using secure scenarios based on reliable historical data~\cite{MaestreEtAl2021}.

Our previous work~\citep{NogueiraEtAl2021} proposed a safe algorithm based on data reconstruction to mitigate the effects of a \fdi{} in the communication between a coordinator and local constraint-free agents, which were bound by global equality constraints.

@@

\todo{
This paper is organized as follows.
First, in Section~\ref{sec:PS}, we introduce the primal decomposition-based DMPC and its vulnerabilities.
Then in Section~\ref{sec:TSM} we study the local problems and propose a detection and mitigation mechanism based on the Expectation Maximization algorithm.
Moreover, in Section~\ref{sec:App_dMPC}, we illustrate the algorithm with academic simulation and we assess its performance.
Finally, in Section~\ref{sec:CC}, we conclude, and we give an outlook of future works.
}

\emph{Notation:}
In this paper, $\norm{\cdot}$, $\norm{\cdot}_{F}$, and $\norm{\vec{v}}_{Y}$ represent the $\ell_{2}$ norm, the Frobenius norm and the weighted norm $\norm{Y^{\frac{1}{2}}\vec{v}}$.
$\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto $\set{T}$.
% $\card{x}$ is the number of elements in $x$.
$\kron{}{}$ ($\hadamard{}{}$) is the Kronecker (Hadamard) product.
$\vectorize{A}$ vectorizes matrix $A$.
${n\mathbin{:}i\mathbin{:}j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\mathbin{:}j}$ is equal to ${n\mathbin{:}1\mathbin{:}j}$.
$[A,B]$ ($[A;B]$) is horizontal (vertical) concatenation of matrices.
${\0_{m}=0_{m\times 1}}$ (${\1_{m}=1_{m\times 1}}$), where
$0_{m\times n}$ ($1_{m\times n}$) is a $0$ ($1$) filled \mbox{$m$-by-$n$} matrix.
$\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
$\preceq$ is the component-wise inequality for vectors and matrix inequality for matrices in ${\symmetric^{n}=\setbuild{X:\R^{n\times n}}{X\T=X}}$.
${\defpos^{n}=\setbuild{X:\symmetric^{n}}{X\succ 0}}$, ${\semidefpos^{n}=\setbuild{X:\symmetric^{n}}{X\succeq 0}}$.
$I_{c}$ is the identity matrix of size $c$.
$A^{\dagger}={{(A^{T}A)}^{-1}A^{T}}$.
$\determinant{A}$ is the determinant of $A$.
A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
$\random{a},\random{A},\randomvec{a}$ are random variables, scalar, matrix and vector respectively.
$\expectation{\random{x}}$ is the expected value of $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
% $\pdf{\cdot}$ is a probability density function.
${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.

@@

\subsection{Model Predictive Control}\label{ssec:MPC}
Consider a system composed of discrete-time linear time-invariant agents ${i\in\set{M}=\{1\mathbin{:}\nsubsystems\}}$, modeled by
\begin{equation}
\begin{matrix}
  \label{eq:systems}
\vec{x}_{i}[k+1]&=&A_{i}\xik &+& B_{i}\uik\\
\vec{y}_{i}[k+1]&=&C_{i}\xik &&
\end{matrix},
\end{equation}
with $\vec{x}_{i}[k]:\R^{n_{x}}$, $\vec{u}_{i}[k]:\R^{n_{u}}$ and $\vec{y}_{i}[k]:\R^{n_{y}}$.

Each agent's input $\uik$ is constrained by
\begin{equation}
  \label{eq:u_local_constraints}
  \0_{n_{u}}\preceq\uik\preceq\vec{u}_{\max},
\end{equation}
with ${\umax:\R^{n_{u}}}$.
The agents are coupled by global input constraints with weighting matrices \todo{${\Gamma_{i}:\semidefpos^{n_{u}}}$}:
\begin{equation}
  \label{eq:global_constraint}
  \sum_{i\in\set{M}}\Gamma_{i}\uik\preceq\umax.
\end{equation}

The overall system is controlled by a \mpc{}, which
computes the optimal input for a finite prediction horizon ${\set{H}=\{1\mathbin{:}\predhorz\}}$ by solving the following problem:
\begin{equation}\label{eq:GOP}
\resizebox{.88\columnwidth}{!}{$
\begin{matrix}
\underset{\useq}{\mathrm{minimize}}&\resizebox{0.35\textwidth}{!}{$\overbrace{\sum\limits_{i\in\set{M}} \overbrace{\sum_{j\in\set{H}}\norm{\mpcvec{v}[i][k+j][k]}^{2}_{Q_i}+\norm{\mpcvec{u}[i][k+j-1][k]}^{2}_{R_i}}^{\textstyle{} \obji}}^{\textstyle{} \globobj}$}\\
\mathrm{subject~ to}&~\eqref{eq:systems},\eqref{eq:u_local_constraints}\ \mathrm{and}\ \eqref{eq:global_constraint}
\left\}\small
\begin{aligned}
  &\forall i\in\set{M}\\
  &\forall j\in\set{H}
\end{aligned}\right.,

@@

Solving~\eqref{eq:GOP} at each time $k$ can be quite challenging, since the prediction horizon and the number of agents can be large making the process computationally intensive. Another issue that can arise is the fact that the complete information \todo[verify]{(${\set{I}_{i}=\{A_{i},B_{i},C_{i},\vec{v}_{i}[k],Q_{i},R_{i}\}}$)} from all the subsystems is needed to solve the problem,
which can be viewed negatively in a confidentiality point of view.

So, we use the \emph{primal decomposition} method, sometimes called \emph{quantity decomposition} or \emph{resource allocation}, with wich we can profit from the parallelism, while avoiding the use of the complete information $\set{I}_{i}$.

@@

The technique consists of decomposing the monolithic \mpc{} in~\eqref{eq:GOP} into $M$ modified \mpc{} problems~\eqref{eq:DOP_local}, called \emph{local problems}, which can be solved in parallel by each subsystem, and a \emph{master problem}~\eqref{eq:DOP_master}, which is equivalent to the global problem and is solved by a coordinator.
\begin{subequations}
  \begin{equation}
    \left.
      \small
        J_{i}^{\star}(\thetaik)=
        \begin{matrix}
        \underset{\uiseq}{\mathrm{minimize}}&\obji\\
        \mathrm{subject~ to} &\eqref{eq:systems}\ \&\ \eqref{eq:u_local_constraints}\\
        &\Gamma_{i}\uik\preceq\qik:\dik\\
      \end{matrix}
    \right\}
    \small
    \begin{aligned}
      &\forall i\in\set{M}\\
      &\forall j\in\set{H}
    \end{aligned}
    \label{eq:DOP_local}
  \end{equation}

  \begin{equation}
    \small
    \begin{aligned}
      J^{\star}=
      \begin{matrix}
        \underset{\qiseq}{\mathrm{minimize}} &\sum_{i\in\set{M}} J^{\star}_i(\qik)\\
        \mathrm{subject~ to} &
        \left.
          \begin{aligned}
            \quad \sum_{i\in\set{M}}\qik\preceq\umax\\
            \quad \qik[k]\succeq\0_{n_{u}},       \forall i\in\set{M}
          \end{aligned}
        \right\}
        \begin{aligned}
          \forall j\in\set{H}
        \end{aligned}
      \end{matrix}
    \end{aligned}
    \label{eq:DOP_master}
  \end{equation}
\end{subequations}
The modified \mpc{} problems in~\eqref{eq:DOP_local} are created by exchanging the $\vec{u}_{\max}$ in~\eqref{eq:global_constraint} by vectors $\qik$, where which one corresponds to the quantity of the total resource $\vec{u}_{\max}$ that is allocated to agent $i$ in time $k$ for each prediction $j$, thus the names \emph{resource allocation} and \emph{quatity decomposition}. This new set of constraints have associated dual variables ${\dik\succeq\0_{n_{u}}}$. The sequence of this variables ${\qiseq}$ and ${\diseq}$ can be aggregated in vectors ${\thetaik=\kron{\1_{\predhorz}}{\qik}}$, and ${\lambdaik=\kron{\1_{\predhorz}}{\dik}}$.

The \emph{master problem} finds the optimal allocation sequences $\optthetai$ by using a projected sub-gradient method:
\begin{equation}
  \label{eq:projectedSubgradient}
\vec{\theta}[k]\pplusone=\Proj^{\set{S}}(\vec{\theta}[k]\p-\rho\p\vec{g}[k]\p),
\end{equation}
where ${\set{S} = \setbuild{\vec{\theta}}{I_{c}^{M}\vec{\theta}\preceq \vec{U}_{\max}\ \&\ \vec{\theta}\succeq\0 }}$, being ${c=\predhorz n_{u}}$, ${I_{c}^{M}=\kron{\1_{M}}{I_{c}}}$, ${\vec{U}_{\max}=\kron{\1_{\predhorz}}{\vec{u}_{\max}}}$, $(p)$ is a given step in the iterative process and $\vec{g}\p[k]$ is a sub-gradient of the objective function of problem in~\eqref{eq:DOP_master} in step $(p)$.

From~\cite{BoydEtAl2015}, it is known that $-\lambdaik$ are sub-gradients of objective $J[k]$.
Plugging it in $\vec{g}[k]\p$ we have
\begin{equation}
  \label{eq:thetaNegot}
\vec{\theta}[k]\pplusone=\Proj^{\set{S}}(\vec{\theta}[k]\p+\rho\p\vec{\lambda}[k]\p),
\end{equation}
\newcommand{\negotiation}{\emph{negotiation}}
which henceforth is referred as the \negotiation.

@@

\begin{algorithm2e}[h]
  \DontPrintSemicolon%
  $p:=0$\;
  Coordinator initializes $\vec{\theta}\p$ \;
  \Repeat{$\|\vec{\theta}^{(p)} -\vec{\theta}^{(p-1)}\|\leq\epsilon$}{
  Subsystems solve~\eqref{eq:DOP_local}, and send $\optlambdai (\thetai\p)$\;
  Coordinator updates allocations~\eqref{eq:thetaNegot}\;
  $p:=p+1$
}
 \caption{Quantity decomposition based \acrshort{dmpc}.}\label{alg:quantityAlg}
\end{algorithm2e}

One can observe that instead of using $\mathcal{I}_{i}$, each agent only sends $\lambdaik$.
An interpretation of $\lambdaik$ is the dissatisfaction of agent $i$ with the resource $\thetaik$ allocated for it, where ${\lambdaik=\0_{c}}$ means total satisfaction.
The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations, but if false data is injected, the \negotiation{} can be driven by a malicious agent which can take advantage of this trust to favor itself, harm others or even destabilize the \negotiation{}, as shown in~\cite{NogueiraEtAl2021}.

Our main objective is to mitigate the effects of a given configuration where agents send untrustworthy
\begin{equation*}
\lambdaicheat=\gamma_{i}(\lambdai).
\end{equation*}

@@

The \emph{local problems}~\eqref{eq:DOP_local} can be rewritten in a equivalent (same solution) constrained \qp{} form:
\todo[Correct alignment maybe]{
  \begin{subequations}\label{eq:quadratic_case}
    \begin{align}
      \label{eq:4}
      \underset{\Uik}{\mathrm{minimize}}&& \overbrace{\frac{1}{2} \Uik^T H_i\Uik+{\fik}^T\Uik}^{\textstyle J_{i}(\vec{\theta}_{i})}\\
      \mathrm{s.t.}&&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\label{eq:constraints}\\
                                        &&\Uik\succeq\0_{c},
    \end{align}
  \end{subequations}
}
where ${\Uik}$ stacks the input prediction sequences, and
${\bar{\Gamma}_{i}=\kron{I_{\predhorz}}{\Gamma_{i}}}$.
The exact values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$. The approach presented in this paper does not depend on the choice of the control objective.
The approach depends only on the fact that $H_{i}:\semidefpos$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.

\no{
  We choose reference tracking just to illustrate the approach. With this objective, we have
${\vik=\wik-C_{i}\xik}$, where $\wik$ is the output reference in time $k$, which yields
\begin{equation}
\small\begin{matrix*}[l]
 H_i&=&\mathcal{D}_i^T\bar{Q}_i\mathcal{D}_i+\bar{R}_i\\
\fik&=&\mathcal{D}_i^T\bar{Q}_i(\mathcal{M}_i\xik-\Wik),
\end{matrix*}
\label{eq:matrices}
\end{equation}
where $\Wik$ stacks the setpoint prediction sequence, which is supposed to be constant during the predicted horizon ${\Wik=\kron{\1_{\predhorz}}{\vec{w}_{i}[k]}}$ (no prior information).
${\mathcal{M}_{i}}$  and ${\mathcal{D}_{i}}$ are the prediction matrices of the \mpc{}.
${\bar{Q}_{i}=\kron{I_{\predhorz}}{Q_{i}}}$, and ${\bar{R}_{i}=\kron{I_{\predhorz}}{R_{i}}}$.
}
\todo[make a better figure if time permits]{
\begin{figure}[t]
  \centering
  \begin{tikzpicture}[node distance=1.5cm]
    \tikzset{
      agent/.style={draw,circle,minimum width=1cm,},
      coord/.style={draw,circle,minimum width=1cm,very thick}
    }
    % \draw[thick,blue,rounded corners=10pt] (-1.7,1.2) rectangle (1.7,-1.7);
    % \node at (1.2,-1.3) {O};
    % \node (pi) at (-2.5,.5) {$\pi$};
    % \node (phi) at (2.5,-1.) {$\phi_{c}$};

    \node[agent] (a1) {a$_{1}$};
    \node (a2) [right of=a1]  {$\dots$};
    \node[agent] (am) [right of=a2] {a$_{M}$};
    \node[coord] (c) [above of=a2] {C};

    \draw[-latex] (a1) to [out=60,in=210] node [sloped,above] {$\vec{\lambda_{1}}[k]$} (c) ;
    \draw[-latex] (c) to [out=-120,in=30] node [sloped,below] {$\vec{\theta}_{1}[k]$} (a1);

    \draw[-latex] (am) to [out=150,in=-60]node [sloped,below] {$\vec{\lambda}_{M}[k]$} (c);
    \draw[-latex] (c) to [out=-30,in=120] node [sloped,above] {$\vec{\theta}_{M}[k]$} (am);

  \end{tikzpicture}
  \caption{Exchange between coordinator (C) and agents $a_{i}$.}\label{fig:dmpc_graph}
\end{figure}
}

We can get an explicit solution for the dual variables $\vec{\lambda}_{i}[k]$, which are \acrfull{pwa} functions w.r.t. $\vec{\theta}_{i}[k]$:
% \begin{equation}
%   \resizebox{.88\columnwidth}{!}{$
%     \begin{aligned}
%       \label{eq:lambdafuntheta}
%       \lambdaik=
%       \begin{cases}
%         -P_{i}^{1}\thetaik-\vec{s}_{i}^{1}[k]&\text{if}\ G^{1}[k]\thetaik \preceq b^{1}[k] \\
%         \qquad\quad \vdots&\qquad\quad \vdots\\
%         -P_{i}^{N}\thetaik-\vec{s}_{i}^{N}[k]&\text{if}\ G^{N}[k]\thetaik \preceq b^{N}[k] \\
%       \end{cases}
%     \end{aligned}
%     $}
% \end{equation}
\begin{equation}
\resizebox{.88\columnwidth}{!}{$
  \begin{aligned}\label{eq:lambdafuntheta}
    \lambdaik=
    % \begin{cases}
      -P_{i}^{n}\thetaik-\vec{s}_{i}^{n}[k]\text{, if}\ G^{n}[k]\thetaik \preceq b^{n}[k],
    % \end{cases}
  \end{aligned}
  $}
\end{equation}
with $n\in\{1\mathbin{:}N\}$.

The halfspaces defined by the set of pairs $(G^{n}[k],b^{n}[k])$ represent the combinations of active constraints~\eqref{eq:constraints} for a given time $k$, which vary w.r.t. $\vec{v}_{i}[k]$.

The $P_{i}^{n}$ are constructed using $H_{i}$ and $\bar{\Gamma}_{i}$; and $\vec{s}_{i}^{j}$ are constructed using $H_{i}$, $\fik$ and $\bar{\Gamma}_{i}$. The specific construction depends on the active constraints in~\eqref{eq:constraints}.
For example, if all constraints are active, we have ${P_{i}^{\text{ac}}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{\text{ac}}[k]=P_{i}^{1}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
And if all constraints are inactive, we have ${P_{i}^{\text{in}}=0_{c\times c}}$ and ${\vec{s}_{i}^{\text{in}}=\0_{c}}$.

\begin{remark}\label{rmk:P_constant}
The $\vec{s}_{i}^{j}[k]$ depend on time $k$, while $P_{i}^{j}$ do not.
\end{remark}
